An Embedding Framework for Consistent Polyhedral Surrogates

### 2019-05-25 neurips-19/poly-embed.tex

\bo{Removed comments to point to appendix, since that creates more links...}
Appendix~\ref{app:calibration} contains the full proof; this section provides a sketch along with the main construction and result.
The first step is to give a link $\psi$ such that exactly minimizing expected surrogate loss $L$, followed by applying $\psi$, always exactly minimizes expected original loss $\ell$.
The existence of such a link is somewhat subtle, because in general some point $u$ that is far from any embedding point can minimize expected loss for two very different distributions $p,p'$, making it unclear whether there exists a link choice $\psi(u)$ that is simultaneously optimal for both.
We show that as we vary $p$ over the infinite $\simplex$, there are only finitely many sets of the form $U = \argmin_{u \in \reals^d} \langle p, L(u) \rangle$.
Associate each $U$ with $R_U \subseteq \R$, the set of reports whose embedding points are in $U$.
Now enforce that all points in $U$ link to some report in $R_U$.
(As a special case, embedding points must link to their corresponding reports.)
Proving this is well-defined uses a chain of arguments involving the Bayes risk, ultimately showing that if $u$ lies in multiple $U$, the corresponding report sets $R_U$ all intersect at some $r =: \psi(u)$.

@@

  \emph{Calibrated:} By construction of the thickened link, if $u$ maps to a wrong report, i.e. $\psi(u) \not\in \gamma(p)$, then $u$ must have at least distance $\epsilon$ to the optimal set $U$.
  We then show that the minimal gradient of the expected loss along any direction away from $U$ is lower-bounded, giving a constant excess expected loss at $u$.
\end{proof}

@@

\begin{figure}
\begin{center}
\begin{minipage}{0.32\linewidth}
\includegraphics[width=\linewidth]{./tikz/abstain-link-linf.pdf}
\end{minipage}\hfill
\begin{minipage}{0.32\linewidth}
\includegraphics[width=\linewidth]{./tikz/abstain-link-U-regions-l1.pdf}
\end{minipage}\hfill
\begin{minipage}{0.32\linewidth}
\includegraphics[width=\linewidth]{./tikz/abstain-link-l1.pdf}
\end{minipage}\hfill
\caption{Constructing links for the abstain surrogate $L_{1/2}$ with $d=2$. The embedding is shown in bold labeled by the corresponding reports. (L) The map $\Psi$ resulting from Theorem~\ref{thm:eps-thick-calibrated} using $\|\cdot\|_\infty$ and $\epsilon = 1/2$, and a possible link $\psi$ which matches eq.~\eqref{eq:abstain-link} from~\cite{ramaswamy2018consistent}.  (M) An illustration of the thickened sets from Definition~\ref{def:eps-thick-link} for two sets $U \in \U$, using $\|\cdot\|_1$ and $\epsilon = 1$. (R) The $\Psi$ and $\psi$ from Theorem~\ref{thm:eps-thick-calibrated} using $\|\cdot\|_1$ and $\epsilon = 1$}
\label{fig:abstain-links}
\end{center}
\end{figure}

@@

In Lemma~\ref{lem:top-k-optimal-corners}\jessiet{in the Appendix... remove ref for camera ready version}, we show that the set $\U = \{u \in \{0,1\}^n : \|u\|_0 \leq k\}$ is always represented among the optimizers of $L'$, meaning for all $p \in \simplex$ we have $\prop{L'}(p) \cap \U \neq \emptyset$.
From Lemma~\ref{lem:loss-restrict}\raft{was Lemma~\ref{lem:top-k-surrogate-embeds}}, then, $L'$ must embed $L'|_\U$, which gives us $\ell':\R'\to\reals^\Y_+$ via the natural embedding from sets $\R' = \{r \subseteq [n] : |r| \leq k\}$ to their indicator vectors.
\begin{align}\label{eq:ell-2}
\ell'(r)_y &= \Ind{y \not\in r} + \tfrac 1 k |r \setminus \{y\}|~.
\end{align}
We now see that $\ell'$ is essentially $\elltopk$ (extended to smaller sets) plus an additional cardinality term.

Knowledge of what loss $L'$ actually embeds greatly simplifies the task of proving inconsistency with $\elltopk$.
Specifically, we see that $\ell'$ allows sets of cardinality strictly less than $k$, so we can look for a distribution making one of these smaller sets optimal.
Writing the expected loss, we have $\inprod{p}{\ell'(r)} = 1 - p(r) + \tfrac 1 k (|r|-p(r)) = 1 + \tfrac 1 k |r| - (1 + \tfrac 1 k) p(r)$, where $p(r) = \sum_{i\in r} p_i$.
% It is now easy to see that the optimal $r$ for any fixed cardinality $|r|=c$ is the $c$ most likely labels.
So let us ask when it is better to drop an element $i$ from $r$:
% If $i$ is the index of the $c$th most likely label (so $p_i = p_{[c]}$)
we have $\inprod{p}{\ell'(r)-\ell'(r\setminus\{i\})} = \tfrac 1 k - p_i$, meaning we will drop elements from $r$ until they all have weight at least $\tfrac 1 k$.
In particular, for distributions close to uniform, $r=\emptyset$ is optimal, already giving us inconsistency\raft{If space (nope) we should spell this out like Jessie did: however ties are broken at $u=0$, pick a distribution with bumps going against the ties, and $u=0$ is still optimal.}.
More generally, as $k < n$, the set $P = \{p\in\simplex : \max_{r\in\R'} p(r) < n/k\}$ is full-dimensional and guarantees that at least one of the top $k$ labels has probability $< \tfrac 1 k$.

@@

Let us consider $g_\emptyset = 0$, $g_1 = g_2 = g_{1,2} = 1$, for which $\ell^g$ is merely 0-1 loss on $\Y$.
For consistency, then, for any distribution $p\in\simplex$, we must have that whenever $u \in \argmin_{u'\in\reals^2} p\cdot L^g(u')$, then $\psi(u)$ must be the most likely outcome, in $\argmax_{y\in\Y} p(y)$.
Simplifying eq.~\eqref{eq:lovasz-hinge}, however, we have
\begin{equation}
  \label{eq:lovasz-hinge-abstain}
  L^g(u)_y = \max\bigl\{(1-u_1y_1)_+,(1-u_2y_2)_+\bigr\} = \max\bigl\{1-u_1y_1,1-u_2y_2,0\bigr\}~,
\end{equation}
which is exactly the abstain surrogate~\eqref{eq:abstain-surrogate}.
We immediately conclude that $L^g$ cannot be consistent with $\ell^g$, as the origin will be the unique optimal report for $L^g$ under distributions with $p_y < 0.5$ for all $y$, and one can simply take a distribution which disagrees with the way ties are broken in $\psi$.
For example, if we take $\sgn(0) = 1$, then under $p((1,1)) = p((1,-1)) = p((-1,1)) = 0.2$ and $p((-1,-1)) = 0.4$, we have $\{0\} = \argmin_{u\in\reals^2} p\cdot L^g(u)$, yet $\psi(0) = (1,1) \notin \{(-1,-1)\} = \argmin_{r\in\R} p\cdot\ell^g(r)$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion and Future Directions} \label{sec:conclusion}
% \paragraph{Summary.}
This paper formalizes an intuitive way to design convex surrogate losses for classification-like problems -- by embedding the reports into $\reals^d$.
We establish a close relationship between embedding and polyhedral surrogates, showing both that every polyhedral loss embeds a discrete loss (Theorem~\ref{thm:poly-embeds-discrete}) and that every discrete loss is embedded by some polyhedral loss (Theorem~\ref{thm:discrete-loss-poly-embeddable}).
We then construct a calibrated link function from a surrogate to the discrete loss it embeds, showing the consistency of our surrogates.
We conclude with examples of how the embedding framework presented can be applied to understand surrogates given in the literature, including those for the abstain loss, top-$k$ loss, and Lov\'asz Hinge.\bo{commented stuff for space}
%We use our link construction to understand calibrated links for abstain different than the one given by~\citet{ramaswamy2018consistent}.
%While one top-$k$ surrogate has already been shown to be inconsistent, we show the similar discrete loss that the surrogate embeds.
%Finally, we use our framework to prove the inconsistency of the Lov\'asz Hinge.

@@

% \paragraph{Directions.}
One open question of particular interest involves the dimension of the input to a surrogate; given a discrete loss, can we construct the surrogate that embeds it \emph{of minimal dimension}?
Often, we na\"ively embed the reports into an $n$-dimensional space, so the dimensionality of the problem increases with the number of possible labels.
As the dimension of the optimization problem is a function of this \emph{embedding dimension},  we would like to understand when it can be reduced.
%There are several direct open questions involving the dimension of the surrogate loss.
%It would be interesting and perhaps practically useful to develop further techniques for upper bounds: automatically constructing embeddings in $\reals^d$ and associated polyhedral losses from a given property, with $d$ as small as possible.
%Another direction is additional lower bound techniques, or further development of our necessary conditions.

@@

\raf{ACKNOWLEDGEMENTS (commented out for submission)}
% \subsection*{Acknowledgements}
% We thank Peter Bartlett for several discussions early on, which led to a proof of \raf{1-d reduction} among other insights.
% Arpit
% Eric Balkanski for a proof of \raf{Lemma on submodularity}
% NSF grant
\bibliographystyle{plainnat}
\bibliography{diss,extra}

>>>>

\bo{Removed comments to point to appendix, since that creates more links...}
Appendix~\ref{app:calibration} contains the full proof; this section provides a sketch along with the main construction and result.
The first step is to give a link $\psi$ such that exactly minimizing expected surrogate loss $L$, followed by applying $\psi$, always exactly minimizes expected original loss $\ell$.
The existence of such a link is somewhat subtle, because in general some point $u$ that is far from any embedding point can minimize expected loss for two very different distributions $p,p'$, making it unclear whether there exists a link choice $\psi(u)$ that is simultaneously optimal for both.
We show that as we vary $p$ over $\simplex$, there are only finitely many sets of the form $U = \argmin_{u \in \reals^d} \inprod{p}{L(u)}$ (Lemma~\ref{lem:polyhedral-range-gamma}).
Associating each $U$ with $R_U \subseteq \R$, the set of reports whose embedding points are in $U$, we enforce that all points in $U$ link to some report in $R_U$.
(As a special case, embedding points must link to their corresponding reports.)
Proving this is well-defined uses a chain of arguments involving the Bayes risk, ultimately showing that if $u$ lies in multiple $U$, the corresponding report sets $R_U$ all intersect at some $r =: \psi(u)$.

@@

  \emph{Calibrated:} By construction of the thickened link, if $u$ maps to an incorrect report, i.e.\ $\psi(u) \not\in \gamma(p)$, then $u$ must have at least distance $\epsilon$ to the optimal set $U$.
  We then show that the minimal gradient of the expected loss along any direction away from $U$ is lower-bounded, giving a constant excess expected loss at $u$.
\end{proof}

@@

\begin{figure}
\begin{center}
\begin{minipage}{0.32\linewidth}
\includegraphics[width=\linewidth]{./tikz/abstain-link-linf.pdf}
\end{minipage}\hfill
\begin{minipage}{0.32\linewidth}
\includegraphics[width=\linewidth]{./tikz/abstain-link-U-regions-l1.pdf}
\end{minipage}\hfill
\begin{minipage}{0.32\linewidth}
\includegraphics[width=\linewidth]{./tikz/abstain-link-l1.pdf}
\end{minipage}\hfill
\caption{Constructing links for the abstain surrogate $L_{1/2}$ with $d=2$. The embedding is shown in bold labeled by the corresponding reports. (L) The link envelope $\Psi$ resulting from Theorem~\ref{thm:eps-thick-calibrated} using $\|\cdot\|_\infty$ and $\epsilon = 1/2$, and a possible link $\psi$ which matches eq.~\eqref{eq:abstain-link} from~\cite{ramaswamy2018consistent}.  (M) An illustration of the thickened sets from Definition~\ref{def:eps-thick-link} for two sets $U \in \U$, using $\|\cdot\|_1$ and $\epsilon = 1$. (R) The $\Psi$ and $\psi$ from Theorem~\ref{thm:eps-thick-calibrated} using $\|\cdot\|_1$ and $\epsilon = 1$.}
\label{fig:abstain-links}
\end{center}
\end{figure}

@@

In Lemma~\ref{lem:top-k-optimal-corners}\jessiet{in the Appendix... remove ref for camera ready version}, we show that the set $\U = \{u \in \{0,1\}^n : \|u\|_0 \leq k\}$ is always represented among the optimizers of $L'$, meaning for all $p \in \simplex$ we have $\prop{L'}(p) \cap \U \neq \emptyset$.
From Lemma~\ref{lem:loss-restrict}\raft{was Lemma~\ref{lem:top-k-surrogate-embeds}}, then, $L'$ must embed $L'|_\U$, which gives us $\ell':\R'\to\reals^\Y_+$ via the natural embedding from sets $\R' = \{r \subseteq [n] : |r| \leq k\}$ to their indicator vectors:
\begin{align}\label{eq:ell-2}
\ell'(r)_y &= \Ind{y \not\in r} + \tfrac 1 k |r \setminus \{y\}| = (1+\tfrac 1 k)\Ind{y \not\in r} + \tfrac 1 k (|r|-1)~.
\end{align}
We now see that $\ell'$ is essentially $\elltopk$ (extended to sets smaller than $k$) plus an additional cardinality term which rewards smaller sets.

Knowledge of what loss $L'$ actually embeds greatly simplifies the task of proving inconsistency with $\elltopk$.
Specifically, we see that $\ell'$ allows sets of cardinality strictly less than $k$, so we can look for a distribution making one of these smaller sets optimal.
Writing the expected loss, we have
%$\inprod{p}{\ell'(r)} = 1 - p(r) + \tfrac 1 k (|r|-p(r)) = 1 + \tfrac 1 k |r| - (1 + \tfrac 1 k) p(r)$,\\
$\inprod{p}{\ell'(r)} = (1+\tfrac 1 k)(1 - p(r)) + \tfrac 1 k (|r|-1) = 1 + \tfrac 1 k |r| - (1 + \tfrac 1 k) p(r)$, where $p(r) = \sum_{i\in r} p_i$.
% It is now easy to see that the optimal $r$ for any fixed cardinality $|r|=c$ is the $c$ most likely labels.
So let us ask when it is better to drop an element $i$ from $r$:
% If $i$ is the index of the $c$th most likely label (so $p_i = p_{[c]}$)
we have $\inprod{p}{\ell'(r)-\ell'(r\setminus\{i\})} = \tfrac 1 k - (1+\tfrac 1 k)p_i$, meaning we will drop elements from $r$ until they all have weight at least $\tfrac 1 {k+1}$.
In particular, for distributions close to uniform, $r=\emptyset$ is optimal, already giving us inconsistency\raft{If space (nope) we should spell this out like Jessie did: however ties are broken at $u=0$, pick a distribution with bumps going against the ties, and $u=0$ is still optimal.}.
More generally, as $k < n$, the set $P = \{p\in\simplex : \max_{r\in\R'} p(r) < k/(k+1)\}$ is full-dimensional and guarantees that at least one of the top $k$ labels has probability strictly less than $\tfrac 1 {k+1}$.

@@

Let us consider $g_\emptyset = 0$, $g_1 = g_2 = g_{1,2} = 1$, for which $\ell^g$ is merely 0-1 loss on $\Y$.
For consistency, then, for any distribution $p\in\simplex$, we must have that whenever $u \in \argmin_{u'\in\reals^2} p\cdot L^g(u')$, then $\psi(u)$ must be the most likely outcome, in $\argmax_{y\in\Y} p(y)$.
Simplifying eq.~\eqref{eq:lovasz-hinge}, however, we have
\begin{equation}
  \label{eq:lovasz-hinge-abstain}
  L^g(u)_y = \max\bigl\{(1-u_1y_1)_+,(1-u_2y_2)_+\bigr\} = \max\bigl\{1-u_1y_1,1-u_2y_2,0\bigr\}~,
\end{equation}
which is exactly the abstain surrogate~\eqref{eq:abstain-surrogate} for $d=2$.
We immediately conclude that $L^g$ cannot be consistent with $\ell^g$, as the origin will be the unique optimal report for $L^g$ under distributions with $p_y < 0.5$ for all $y$, and one can simply take a distribution which disagrees with the way ties are broken in $\psi$.
For example, if we take $\sgn(0) = 1$, then under $p((1,1)) = p((1,-1)) = p((-1,1)) = 0.2$ and $p((-1,-1)) = 0.4$, we have $\{0\} = \argmin_{u\in\reals^2} p\cdot L^g(u)$, yet $\psi(0) = (1,1) \notin \{(-1,-1)\} = \argmin_{r\in\R} p\cdot\ell^g(r)$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion and Future Directions} \label{sec:conclusion}
% \paragraph{Summary.}
This paper formalizes an intuitive way to design convex surrogate losses for classification-like problems---by embedding the reports into $\reals^d$.
We establish a close relationship between embedding and polyhedral surrogates, showing both that every polyhedral loss embeds a discrete loss (Theorem~\ref{thm:poly-embeds-discrete}) and that every discrete loss is embedded by some polyhedral loss (Theorem~\ref{thm:discrete-loss-poly-embeddable}).
We then construct a calibrated link function from any polyhedral loss to the discrete loss it embeds, giving consistency for all such losses.
We conclude with examples of how the embedding framework presented can be applied to understand existing surrogates in the literature, including those for the abstain loss, top-$k$ loss, and Lov\'asz hinge.\bo{commented stuff for space}
%We use our link construction to understand calibrated links for abstain different than the one given by~\citet{ramaswamy2018consistent}.
%While one top-$k$ surrogate has already been shown to be inconsistent, we show the similar discrete loss that the surrogate embeds.
%Finally, we use our framework to prove the inconsistency of the Lov\'asz Hinge.

@@

% \paragraph{Directions.}
One open question of particular interest involves the dimension of the input to a surrogate; given a discrete loss, can we construct the surrogate that embeds it \emph{of minimal dimension}?
Often, we na\"ively embed the reports into an $n$-dimensional space, so the dimensionality of the problem increases with the number of possible labels.
As the dimension of the optimization problem is a function of this \emph{embedding dimension}, a promising direction is to leverage tools from elicitation complexity~\cite{lambert2008eliciting,frongillo2015elicitation} and convex calibration dimension~\cite{ramaswamy2016convex} to understand when it can be reduced.
%There are several direct open questions involving the dimension of the surrogate loss.
%It would be interesting and perhaps practically useful to develop further techniques for upper bounds: automatically constructing embeddings in $\reals^d$ and associated polyhedral losses from a given property, with $d$ as small as possible.
%Another direction is additional lower bound techniques, or further development of our necessary conditions.

@@

\raf{ACKNOWLEDGEMENTS (commented out for submission)}
% \subsection*{Acknowledgements}
% We thank Peter Bartlett for several discussions early on, which led to a proof of \raf{1-d reduction} among other insights.
% Arpit
% Eric Balkanski for a proof of \raf{Lemma on submodularity}
% NSF grant
\newpage
\bibliographystyle{plainnat}
\bibliography{diss,extra}
